Practical implications of HATEOAS on a bridge between LOD and SPARQL

### 2012-02-29 eswc-2012/workshop.tex
more headers
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@127 43ef1190-90fd-4dd7-a769-d0121f02b80e
radical RESTafarianism
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@128 43ef1190-90fd-4dd7-a769-d0121f02b80e

\section{Finding the way with VoID}

<> void:inDataset [ void:sparqlEndpoint </sparql> . ] .

@@

>>>>

\section{Proposed solutions}

Prefixes are in the following omitted for brevity.

\subsection{Simple hypermedia vocabulary}

To help drive interaction through hypermedia, the RDF information
resources has to contain information about the URIs that can be used
to perform operations other than pure reads. To this end, we suggest a
simple vocabulary that can be used to supply some statements with all
resources that the client may be allowed to manipulate.

In keeping with the REST philosophy that the representation should be
independent of the protocol that is used to transfer it, the
vocabulary is independent of HTTP, but the explanatory comments uses
HTTP examples, as this makes the discussion clearer.

\begin{verbatim}
hm:operations a rdf:Property ;
              rdfs:range foaf:Document ;
              rdfs:domain hm:Operations .
\end{verbatim}
\texttt{hm:operations} link the resource we want to manipulate,
typically the current information resource to the operations we may be
allowed to perform on it.

In all the following examples, the object of the statement is the
Request-URI that the server suggests is used to manipulate the resource.
\begin{verbatim}
hm:mergeInto a rdf:Property ;
             rdfs:range hm:Operations ;
             rdfs:domain foaf:Document .
\end{verbatim}
\texttt{hm:mergeInto} is used to indicate that a HTTP POST operation
will do an RDF merge of the request body with the content at the
Request-URI.

\begin{verbatim}
hm:serverCreate a rdf:Property ;
                rdfs:range hm:Operations ;
                rdfs:domain foaf:Document .
\end{verbatim}
\texttt{hm:serverCreate} is used to indicate that a HTTP POST
operation to the Request-URI will create a new resource with a
server-supplied URI.

\begin{verbatim}
hm:delete a rdf:Property ;
          rdfs:range hm:Operations ;
          rdfs:domain foaf:Document .
\end{verbatim}
\texttt{hm:delete} is used to indicate that the resource should be
deleted when issuing a HTTP DELETE operation to the Request-URI.

\begin{verbatim}
hm:put a rdf:Property ;
       rdfs:range hm:Operations ;
       rdfs:domain foaf:Document .
\end{verbatim}
\texttt{hm:put} is used to indicate that the resource should be
created using the request body or updated with it if it already exists
when issuing a HTTP PUT operation to the Request-URI.

\begin{verbatim}
hm:Operations a rdfs:Class .
\end{verbatim}
The class of possible operations.

To manipulate default graphs, properties \texttt{hm:mergeIntoDefault},
\texttt{hm:deleteDefault} and \texttt{hm:putDefault} are defined
similarly.

In most cases, the current information resource will be the subject of
the \texttt{hm:operations} \strong{but also} the object of the allowed
operations. A typical information resource will supply the following
triples (given a suitable base URI):

\begin{verbatim}
<> hm:operations
   [ a hm:Operations ;
     hm:mergeInto <> ;
     hm:delete <> ;
     hm:put <> ;
     hm:serverCreate <../> .
   ] .
\end{verbatim}
This represents the common case when you GET a resource, then update
it with PUT or DELETE or append triples with POST.

\subsection{Finding the way with VoID}

\begin{verbatim}
<> void:inDataset [ void:sparqlEndpoint </sparql> . ] .
\end{verbatim}

@@

\item What can we do to add more H Factors\footnote{see
  \url{http://amundsen.com/hypermedia/hfactor/}} to RDF?
